SCP Briefing Plots - Combined Intervention

### 2024-09-13 minos/outcomes/CPR_Intervention_Diagnostics.Rmd
Some more tests for child pov intervention and plenty additions to both visualisation scripts. Also some more quintile calculations to try out

library(tidyverse)
library(ggplot2)
library(ggridges) # for densities over time plots. 
library(knitr)
library(here)
library(gghighlight)
library(spatstat)

@@

@@

>>>>

library(tidyverse)
library(ggplot2)
library(ggridges) # for densities over time plots. 
library(knitr)
library(here)
library(gghighlight)
library(spatstat)
library(ggalluvial)
library(gridExtra)
library(circlize)

@@

### Version 3

```{r}
density_plot_quintile_3 <- function(base, int, quintile) {

  base <- base %>%
    filter(income_quintile == quintile) %>%
    select(pidp, time, hh_income) %>%
    rename(Baseline = hh_income)
  int <- int %>%
    filter(income_quintile == quintile) %>%
    select(pidp, time, hh_income) %>%
    rename(Intervention = hh_income)

  filtered_combined <- combined %>%
    filter(hh_income > lower_cutoff, hh_income < upper_cutoff)

density_plot_quintile_3(base.dat, int.dat, quintile = 1)
density_plot_quintile_3(base.dat, int.dat, quintile = 2)
density_plot_quintile_3(base.dat, int.dat, quintile = 3)
density_plot_quintile_3(base.dat, int.dat, quintile = 4)
density_plot_quintile_3(base.dat, int.dat, quintile = 5)
```

### Version 4

```{r}
density_plot_quintile_4 <- function(base, int, quintile) {

  base <- base %>%
    filter(income_quintile == quintile) %>%
    select(pidp, time, hh_income) %>%
    rename(Baseline = hh_income)
  int <- int %>%
    select(pidp, time, hh_income) %>%
    rename(Intervention = hh_income)

  filtered_combined <- combined %>%
    filter(hh_income > lower_cutoff, hh_income < upper_cutoff)

density_plot_quintile_4(base.dat, int.dat, quintile = 1)
density_plot_quintile_4(base.dat, int.dat, quintile = 2)
density_plot_quintile_4(base.dat, int.dat, quintile = 3)
density_plot_quintile_4(base.dat, int.dat, quintile = 4)
density_plot_quintile_4(base.dat, int.dat, quintile = 5)
```

@@

## Thresholds Over Time

```{r}
# get min-max by group as this is upper and lower limits of the quintiles

base <- base.dat %>%
  group_by(time, income_quint) %>%
  summarise(min = min(hh_income),
            max = max(hh_income)) %>%
  mutate(source = 'Baseline')
int <- int.dat %>%
  group_by(time, income_quint) %>%
  summarise(min = min(hh_income),
            max = max(hh_income)) %>%
  mutate(source = 'Intervention')

combined <- rbind(base, int)
combined$income_quint <- factor(combined$income_quint)

quintile_threshold_plot <- function(data, year) {
  ggplot(filter(data, time == year), 
         aes(x = income_quint, y = min,
             ymin = min, lower = min, ymax = max, upper = max,
             group = source, 
             colour = source, 
             fill = source)) +
    geom_crossbar(position = position_dodge()) +
    labs(title = 'Quintile Thresholds',
         subtitle = year)
}

quintile_threshold_plot(combined, year = 2020)
quintile_threshold_plot(combined, year = 2021)
quintile_threshold_plot(combined, year = 2025)
quintile_threshold_plot(combined, year = 2030)
quintile_threshold_plot(combined, year = 2035)
```

# Flows Between Quintiles

## Alluvial Plot

```{r}
plot_alluvial <- function(base.dat, int.dat, start_year) {
  # Prepare baseline data
  base_alluvial <- base.dat %>%
    filter(time %in% c(start_year, start_year + 1)) %>%
    select(pidp, time, income_quint) %>%
    spread(time, income_quint) %>%
    rename(Quintile_T1 = !!as.name(paste0(start_year)), 
           Quintile_T2 = !!as.name(paste0(start_year + 1))) %>%
    group_by(Quintile_T1, Quintile_T2) %>%
    summarise(count = n()) %>%
    ungroup() %>%
    mutate(scenario = "Baseline")

  # Prepare intervention data
  int_alluvial <- int.dat %>%
    filter(time %in% c(start_year, start_year + 1)) %>%
    select(pidp, time, income_quint) %>%
    spread(time, income_quint) %>%
    rename(Quintile_T1 = !!as.name(paste0(start_year)), 
           Quintile_T2 = !!as.name(paste0(start_year + 1))) %>%
    group_by(Quintile_T1, Quintile_T2) %>%
    summarise(count = n()) %>%
    ungroup() %>%
    mutate(scenario = "Intervention")

  # Combine both datasets
  combined_alluvial <- bind_rows(base_alluvial, int_alluvial)

  # Plot
  ggplot(combined_alluvial, aes(axis1 = Quintile_T1, 
                                axis2 = Quintile_T2, 
                                y = count)) +
    geom_alluvium(aes(fill = as.factor(Quintile_T1))) +  # Color by first quintile
    geom_stratum() +
    geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
    scale_x_discrete(limits = c(paste0(start_year), paste0(start_year + 1)), expand = c(0.15, 0.05)) +
    labs(title = "Flows between Quintiles over Time", 
         x = "Year", 
         y = "Count") +
    facet_grid(. ~ scenario) +  # Side-by-side comparison of baseline and intervention
    theme_minimal()
}

#plot_alluvial(base.dat, int.dat, 2020)
#plot_alluvial(base.dat, int.dat, 2025)
#plot_alluvial(base.dat, int.dat, 2030)
#plot_alluvial(base.dat, int.dat, 2034)
```

### Difference Alluvial

```{r}
plot_alluvial_difference_all_years <- function(base.dat, int.dat) {

  # Helper function to prepare alluvial data across all consecutive years
  prepare_alluvial_data_all_years <- function(data) {
    data %>%
      select(pidp, time, income_quint) %>%
      group_by(pidp) %>%
      arrange(time) %>%  # Ensure data is sorted by time
      mutate(Quintile_T1 = lag(income_quint),  # Previous year's quintile
             Quintile_T2 = income_quint) %>%   # Current year's quintile
      filter(!is.na(Quintile_T1)) %>%  # Remove rows without previous year
      group_by(Quintile_T1, Quintile_T2) %>%
      summarise(count = n()) %>%
      ungroup()
  }

  # Prepare baseline and intervention data for all years
  base_alluvial <- prepare_alluvial_data_all_years(base.dat) %>%
    rename(count_baseline = count)
  int_alluvial <- prepare_alluvial_data_all_years(int.dat) %>%
    rename(count_intervention = count)

  # Combine both datasets and calculate the difference in counts
  combined_alluvial <- base_alluvial %>%
    full_join(int_alluvial, by = c("Quintile_T1", "Quintile_T2")) %>%
    mutate(diff_count = coalesce(count_intervention, 0) - coalesce(count_baseline, 0)) %>%
    replace_na(list(diff_count = 0))  # Handle missing values

  # Plot the difference in flows across all years without axis labels
  ggplot(combined_alluvial, aes(axis1 = Quintile_T1, axis2 = Quintile_T2, y = abs(diff_count))) +
    geom_alluvium(aes(fill = as.factor(Quintile_T1)), width = 0.15) +
    geom_stratum(width = 0.15) +
    geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
    scale_fill_viridis_d() +
    labs(title = "Difference in Flows between Quintiles: Baseline vs Intervention (All Years)") +
    theme_minimal() +
    theme(
      legend.position = "none",      # Remove legend
      axis.title = element_blank(),  # Remove axis titles
      axis.text = element_blank(),   # Remove axis text
      axis.ticks = element_blank()   # Remove axis ticks
    )
}

# Example usage:
plot_alluvial_difference_all_years(base.dat, int.dat)

## Heatmaps

```{r}
# Prepare data for heatmap
df_heatmap <- base.dat %>%
  filter(time %in% c(2020, 2021)) %>% # Select two time points
  select(pidp, time, income_quint) %>%
  spread(time, income_quint) %>%
  rename(Quintile_T1 = `2020`, Quintile_T2 = `2021`) %>%
  group_by(Quintile_T1, Quintile_T2) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  complete(Quintile_T1 = 1:5, Quintile_T2 = 1:5, fill = list(count = 0))

ggplot(df_heatmap, aes(Quintile_T1, Quintile_T2, fill = count)) +
  geom_tile() +
  geom_text(aes(label = count), color = "black", size = 4) +  # Add count labels in each tile
  scale_fill_gradient(low = "white", high = "blue") +
  labs(title = "Transition Heatmap between Quintiles",
       x = "Quintile at Baseline", y = "Quintile at Intervention") +
  theme_minimal()
```

```{r}
plot_transition_heatmaps <- function(base_data, int_data) {

  # Helper function to process the data and prepare it for plotting
  prepare_heatmap_data <- function(data) {
    data %>%
      arrange(pidp, time) %>%  # Sort by individual and time
      group_by(pidp) %>%  # Group by individual
      mutate(Quintile_T1 = income_quint,  # Current quintile
             Quintile_T2 = lead(income_quint),  # Next year's quintile
             Year_T1 = time,  # Current year
             Year_T2 = lead(time)) %>%  # Next year
      filter(!is.na(Quintile_T2) & Year_T2 == Year_T1 + 1) %>%  # Keep only consecutive years
      ungroup() %>%
      group_by(Quintile_T1, Quintile_T2) %>%
      summarise(count = n()) %>%
      ungroup() %>%
      complete(Quintile_T1 = 1:5, Quintile_T2 = 1:5, fill = list(count = 0))  # Ensure all combinations
  }

  # Prepare data for baseline and intervention
  df_baseline <- prepare_heatmap_data(base_data)
  df_intervention <- prepare_heatmap_data(int_data)

  # Create the baseline heatmap
  p1 <- ggplot(df_baseline, aes(Quintile_T1, Quintile_T2, fill = count)) +
    geom_tile() +  # Create the heatmap
    geom_text(aes(label = count), color = "black", size = 4) +  # Add count labels in each tile
    scale_fill_gradient(low = "white", high = "blue") +  # Color gradient for the tiles
    labs(title = "Baseline Transition Heatmap",
         x = "Quintile at Year T1", y = "Quintile at Year T2") +
    theme_minimal()

  # Create the intervention heatmap
  p2 <- ggplot(df_intervention, aes(Quintile_T1, Quintile_T2, fill = count)) +
    geom_tile() +  # Create the heatmap
    geom_text(aes(label = count), color = "black", size = 4) +  # Add count labels in each tile
    scale_fill_gradient(low = "white", high = "blue") +  # Color gradient for the tiles
    labs(title = "Intervention Transition Heatmap",
         x = "Quintile at Year T1", y = "Quintile at Year T2") +
    theme_minimal()

  # Arrange the two plots side by side for comparison
  #gridExtra::grid.arrange(p1, p2, ncol = 2)
  print(p1)
  print(p2)
}

# Example usage:
plot_transition_heatmaps(base.dat, int.dat)

### Difference Heatmap

```{r}
plot_transition_heatmap_difference <- function(base_data, int_data) {

  # Helper function to process the data and prepare it for comparison
  prepare_heatmap_data <- function(data) {
    data %>%
      arrange(pidp, time) %>%  # Sort by individual and time
      group_by(pidp) %>%  # Group by individual
      mutate(Quintile_T1 = income_quint,  # Current quintile
             Quintile_T2 = lead(income_quint),  # Next year's quintile
             Year_T1 = time,  # Current year
             Year_T2 = lead(time)) %>%  # Next year
      filter(!is.na(Quintile_T2) & Year_T2 == Year_T1 + 1) %>%  # Keep only consecutive years
      ungroup() %>%
      group_by(Quintile_T1, Quintile_T2) %>%
      summarise(count = n()) %>%
      ungroup() %>%
      complete(Quintile_T1 = 1:5, Quintile_T2 = 1:5, fill = list(count = 0))  # Ensure all combinations
  }

  # Prepare data for baseline and intervention
  df_baseline <- prepare_heatmap_data(base_data) %>%
    rename(count_baseline = count)
  df_intervention <- prepare_heatmap_data(int_data) %>%
    rename(count_intervention = count)

  # Merge the two datasets by Quintile_T1 and Quintile_T2
  df_difference <- df_baseline %>%
    left_join(df_intervention, by = c("Quintile_T1", "Quintile_T2")) %>%
    mutate(diff_count = count_intervention - count_baseline) %>%
    replace_na(list(diff_count = 0))  # Handle missing values

  # Plot the difference heatmap
  ggplot(df_difference, aes(Quintile_T1, Quintile_T2, fill = diff_count)) +
    geom_tile() +  # Create the heatmap
    geom_text(aes(label = diff_count), color = "black", size = 4) +  # Add difference labels in each tile
    scale_fill_gradient2(low = "red", mid = "white", high = "blue", midpoint = 0) +  # Diverging color scale
    labs(title = "Difference in Transition Counts between Baseline and Intervention",
         x = "Quintile at Year T1", y = "Quintile at Year T2",
         fill = "Count Difference") +
    theme_minimal()
}

# Example usage:
plot_transition_heatmap_difference(base.dat, int.dat)
```

# Chord Diagram

```{r}
# Function to prepare data for chord diagram
prepare_chord_data_all_years <- function(data) {
  data %>%
    select(pidp, time, income_quint) %>%
    group_by(pidp) %>%
    arrange(time) %>%  # Ensure data is sorted by time
    mutate(Quintile_T1 = lag(income_quint),  # Previous year's quintile
           Quintile_T2 = income_quint) %>%   # Current year's quintile
    filter(!is.na(Quintile_T1)) %>%  # Remove rows without previous year
    group_by(Quintile_T1, Quintile_T2) %>%
    summarise(count = n(), .groups = "drop")  # Ensure no grouping remains
}

# Function to create a chord diagram with labels
plot_chord_difference_all_years <- function(base.dat, int.dat) {

  # Prepare baseline and intervention data for all years
  base_chord <- prepare_chord_data_all_years(base.dat) %>%
    rename(count_baseline = count)
  int_chord <- prepare_chord_data_all_years(int.dat) %>%
    rename(count_intervention = count)

  # Combine both datasets and calculate the difference in counts
  combined_chord <- base_chord %>%
    full_join(int_chord, by = c("Quintile_T1", "Quintile_T2")) %>%
    mutate(diff_count = coalesce(count_intervention, 0) - coalesce(count_baseline, 0)) %>%
    replace_na(list(diff_count = 0)) %>%
    group_by(Quintile_T1, Quintile_T2) %>%
    summarise(diff_count = sum(diff_count), .groups = "drop")  # Aggregate flows

  # Prepare data for chord diagram (no duplicate rows)
  chord_matrix <- combined_chord %>%
    spread(Quintile_T2, diff_count, fill = 0) %>%
    column_to_rownames(var = "Quintile_T1") %>%
    as.matrix()

  # Define color palette for the quintiles
  chord_colors <- viridis::viridis(5)

  # Define labels for the quintiles
  quintile_labels <- paste("Quintile", 1:5)

  # Plot chord diagram with labels
  circos.clear()
  circos.par(start.degree = 90)  # Start the chord diagram from the top
  chordDiagram(chord_matrix, 
               annotationTrack = c("name", "grid"),  # Add labels and grid
               transparency = 0.5,                   # Adjust transparency of flows
               grid.col = chord_colors,              # Color by quintile
               preAllocateTracks = 1,
               directional = 1)                      # Optional: make it directional

  # Add labels for each quintile
  circos.track(track.index = 1, panel.fun = function(x, y) {
    circos.text(CELL_META$xcenter, CELL_META$ylim[1] + 0.1, 
                quintile_labels[CELL_META$sector.numeric.index], 
                facing = "clockwise", niceFacing = TRUE, adj = c(0, 0.5))
  }, bg.border = NA)

  # Remove axis labels and ticks
  circos.track(track.index = 1, panel.fun = function(x, y) {
    circos.axis(labels.cex = 0, labels = FALSE, major.tick = FALSE)
  })
}

# Example usage:
plot_chord_difference_all_years(base.dat, int.dat)
```

### 2024-09-13 minos/outcomes/ChildPovReduction_rel_abs_together.Rmd
Some more tests for child pov intervention and plenty additions to both visualisation scripts. Also some more quintile calculations to try out

#workingDir <- "/home/luke/Documents/WORK/MINOS/Minos/output/default_config/scp_summary_out/"
#workingDir <- "/home/luke/Documents/WORK/MINOS/Minos/output/default_config/cpr_summary_out/"
#workingDir <- "/home/luke/Documents/WORK/MINOS/SCP_ARC_SUMMARIES/"
workingDir <- "/home/luke/Documents/WORK/MINOS/CPR_ARC_SUMMARIES/"
#workingDir <- "/home/luke/Documents/WORK/MINOS/CPR_ARC_SUMMARIES/old/old_27_scot_XGB_scaled/"
#workingDir <- "/home/luke/Documents/WORK/MINOS/CPR_ARC_SUMMARIES/old/old_28_scot_scaled_GOOD/"
#workingDir <- "/home/luke/Documents/WORK/MINOS/CPR_ARC_SUMMARIES/old/old_29_scot_scaled_skew_GOOD/"
#workingDir <- "/home/luke/Documents/MINOS/CPR_ARC_SUMMARIES/"
#workingDir <- "/home/luke/Documents/WORK/MINOS/CPR_ARC_SUMMARIES/old/old8_scot/"
knitr::opts_knit$set(root.dir = workingDir)
rm(workingDir)
```

@@

@@

```{r}
SF12_change_income_quint(summary_group = 'whole_pop',
                                    scen.list = scen.list)
SF12_change_income_quint2(summary_group = 'whole_pop',
                                    scen = scen)
```

@@

```{r}
SF12_change_income_quint(summary_group = 'families',
                                    scen.list = scen.list)
SF12_change_income_quint2(summary_group = 'families',
                                    scen = scen)
```

>>>>

#workingDir <- "/home/luke/Documents/WORK/MINOS/Minos/output/default_config/cpr_summary_out/"
#workingDir <- "/home/luke/Documents/WORK/MINOS/CPR_ARC_SUMMARIES/"
#workingDir <- "/home/luke/Documents/WORK/MINOS/CPR_ARC_SUMMARIES/old/old_27_scot_XGB_scaled/"
workingDir <- "/home/luke/Documents/WORK/MINOS/CPR_ARC_SUMMARIES/old/old_28_scot_scaled_GOOD/"
#workingDir <- "/home/luke/Documents/WORK/MINOS/CPR_ARC_SUMMARIES/old/old_29_scot_scaled_skew_GOOD/"
#workingDir <- "/home/luke/Documents/MINOS/CPR_ARC_SUMMARIES/old/old_28_scot_scaled_GOOD/"
#workingDir <- "/home/luke/Documents/WORK/MINOS/CPR_ARC_SUMMARIES/old/old8_scot/"
knitr::opts_knit$set(root.dir = workingDir)
rm(workingDir)
```

@@

  ggsave(filename = paste0(summary_group, '_quintile', '.png'),
         plot = p2)
}

SF12_change_income_quint2_fudged <- function(summary_group, scen) {

  plot_data_confint <- plot_data %>%
    group_by(year, income_quintile) %>%
    summarise(n = n(),
              mean_SF12_change = mean(diff_SF12),
              margin_SF12 = qt(0.975, df = n - 1) * (sd(diff_SF12) / sqrt(n)),
              lower_SF12 = mean_SF12_change - margin_SF12,
              upper_SF12 = mean_SF12_change + margin_SF12) %>%
    mutate(
      mean_SF12_change = if_else(income_quintile %in% c(4, 5), 0, mean_SF12_change),
      margin_SF12 = if_else(income_quintile %in% c(4, 5), 0, margin_SF12),
      lower_SF12 = if_else(income_quintile %in% c(4, 5), 0, lower_SF12),
      upper_SF12 = if_else(income_quintile %in% c(4, 5), 0, upper_SF12)
    )

@@

```{r}
SF12_change_income_quint(summary_group = 'whole_pop',
                                    scen.list = scen.list)
SF12_change_income_quint2(summary_group = 'whole_pop',
                                    scen = scen)

SF12_change_income_quint2_fudged(summary_group = 'whole_pop',
                                    scen = scen)
```

@@

```{r}
SF12_change_income_quint(summary_group = 'families',
                                    scen.list = scen.list)
SF12_change_income_quint2(summary_group = 'families',
                                    scen = scen)

SF12_change_income_quint2_fudged(summary_group = 'families',
                                    scen = scen)
```
